% Copyright 2013 Thomas Koenig, Alexander Michel, Michael Baumgart
%
% This file is part of NumericPlots.
%
% NumericPlots is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% any later version.
%
% NumericPlots is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with NumericPlots.  If not, see <http://www.gnu.org/licenses/>.


\documentclass[parskip]{scrartcl}

\usepackage[ansinew]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{amsmath}
\usepackage{siunitx}
% \usepackage{floatbarrier}
% \usepackage{showexpl}
\usepackage{etex} % use e-TeX’s extended register set

% options for package:
% BW: black and white
% AxisStyle=<style>: set standard axis style ([Boxed], Frame, FrameI, None)
\usepackage[
	%BW,
  	xAxisStyle=Lower,
 	yAxisStyle=Left,
 	]{NumericPlots}
 
\usepackage[
	ps2pdf,							%% similar to dvips but redefines some macros for use with ps2pdf
	final,
	bookmarks=true,				%% if true, generate PDF bookmarks (requires two passes of pdflatex)
	bookmarksopen=false,			%% if true, show all PDF bookmarks expanded
	bookmarksnumbered=true,		%% if true, add the section numbers to the bookmarks
% 		linkbordercolor=acin_red,
% 		urlbordercolor=acin_red,
% 		citebordercolor=acin_red,
% 		filebordercolor=acin_red,
	pdfpagemode=UseOutlines, %% None, UseOutlines (=show bookmarks), UseThumbs (show thumbnails), FullScreen
%	breaklinks,
	pdfauthor = {Thomas Koenig Alexander Michel Michael Baumgart},
	pdftitle = {NumericPlots - plot numeric data with latex},
	pdfsubject = {},
	plainpages=false,
	hyperindex=true,
]{hyperref}

\usepackage{graphicx}

\definecolor{NumericPlotsCommands}{RGB}{0, 102, 153}
\usepackage{listings}
\lstset{language=[latex]tex,tabsize=2,basicstyle=\small\ttfamily,%
	numbers=none, numberstyle=\tiny,%
	breaklines=true, breakatwhitespace=false,%
	emptylines=*1,%
	columns=flexible,%
	keywordstyle=\color{black},commentstyle=\color{gray},%
	keywordstyle={[2]\color{NumericPlotsCommands}},%
	morekeywords={[2]setxAxis, setyAxis, NumericDataPlot, plotxAxis, plotyAxis,%
	NDPput,putExpX,putExpY,%
	putN,putNE,putE,putSE,putS,putSW,putW,putNW,%
	PutTickLabelYaxis,PutTickLabelXaxis,%
	NDPhline, NDPvline, NDPline, NDPhbox, NDPvbox, NDPbox,%
	LegendDefinition, LegLine,%
	StdLabelOption,StdTickLabelOption,%
	multilistplot, testframe}%
	}
 	
\title{NumericPlots - plot numeric data with latex}
\author{Thomas K\"{o}nig, Alexander Michel, Michael Baumgart}
\date{\today}

% \includeonly{TestPlots}
% \includeonly{TechnicalDetails}

% input DefineData here, as the examples from DefineData are used throughout the
% document (includeonly still works).
\input{examples/DefineData}
\input{examples/DataTestRealData}

\begin{document}

% redefining the StdLabelOption and StdTickLabelOption applies to all following
% plots.
% \renewcommand{\StdLabelOption}{\LARGE}
% \renewcommand{\StdTickLabelOption}{\LARGE}

\maketitle

Plotting numeric data is a task which often has to be done for scientific
papers. In \LaTeX, however, it is only possible to include graphics created with
an external program. The pstricks-packages provide many commands to
generate graphics in \LaTeX. To generate simple graphics from numeric data,
however, it is difficult to use. This package provides a simpler interface for
the pstricks-package to plot numeric data.

The authors may be reached at
\href{mailto:numericplots@tikey.de}{numericplots@tikey.de}.

\tableofcontents

\part{Introduction}

Plotting numeric data is a task which often has to be done for scientific
papers. In \LaTeX, however, it is only possible to include graphics created with
an external program. The pstricks-packages provide very many commands to
generate graphics in \LaTeX. To generate simple graphics from numeric data,
however, it is difficult to use. This package provides a simpler interface for
the pstricks-package to plot numeric data.

NumericPlots is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
any later version.

NumericPlots is distributed in the hope that it will be useful,
but WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
GNU General Public License for more details.

You should have received a copy of the GNU General Public License
along with NumericPlots.  If not, see http://www.gnu.org/licenses/.

Copyright 2013 Thomas K\"{o}nig, Alexander Michel, Michael Baumgart


% \part{Using the package} -> in BasicFunctionality

\include{BasicFunctionality}

\include{MultiplePlots}

\input{MatlabSupport}

%\section{package options}

%\section{list of commands and options}

\include{options}

\include{TechnicalDetails}

% \part{Examples}

\include{FurtherExamples}

\include{TestPlots}


% \part{Version}

\include{history}

% \section{Test Calculations}
%
% Teste setkeys
%
% Watch:
%
% \setkeys*[NumericDataPlot]{Axis}{xO=34, xMin=10}
% \setrmkeys[NumericDataPlot]{AxisWait}
% 	\begin{tabular}{ll}
% 	xMin & \NumDataPlotxMin\\
% 	\textbf{xO:} & \NumDataPlotxO \\
% 	\end{tabular}
%
% \setkeys*[NumericDataPlot]{Axis}{xMin=10, xO=34}
% \setrmkeys[NumericDataPlot]{AxisWait}
% 	\begin{tabular}{ll}
% 	xMin & \NumDataPlotxMin\\
% 	\textbf{xO:} & \NumDataPlotxO \\
% 	\end{tabular}
%
% \setkeys*[NumericDataPlot]{Axis}{xMin=10}
% \setrmkeys[NumericDataPlot]{AxisWait}
% 	\begin{tabular}{ll}
% 	xMin & \NumDataPlotxMin\\
% 	\textbf{xO:} & \NumDataPlotxO \\
% 	\end{tabular}
%
% calculation example:

% \setkeys*[NumericDataPlot]{Axis}{xO=34, xMax=20, xMin=10, Dx=2}
% \setrmkeys[NumericDataPlot]{AxisWait}
% 	\begin{tabular}{ll}
% 	xCoordMin & \NumDataPlotxCoordMin\\
% 	xCoordMax & \NumDataPlotxCoordMax \\
% 	xMin: & \NumDataPlotxMin \\
% 	xMax: & \NumDataPlotxMax \\
% 	xCoordMax: & \NumDataPlotxCoordMax \\
% 	xCoordMin: & \NumDataPlotxCoordMin \\
% 	Dx: & \NumDataPlotDx \\
% 	\textit{xNrTicks:} & \xNrTicks \\
% 	%\textit{NumDataPlotdx} & \NumDataPlotdx\\
% 	%\textit{NumDataPlotxCoordOrig} & \NumDataPlotdx\\
% 	\textbf{xO:} & \NumDataPlotxO \\
% 	\end{tabular}

\end{document}
