\documentclass{article}
\usepackage{ran_toks}

%\useThisSeed{606574325}
%\useLastAsSeed
%\rtdebugtrue
%\ranToksOn
%\ranToksOff

\parindent0pt \parskip6pt


\begin{document}
\begin{center}\bfseries
    Test file for \textsf{ran\_toks} Package\\[3pt]
        D. P. Story
\end{center}

Test of the \verb!\ranToks! command.

\ranToks{myPals}{%
    {Jim}{Richard}{Don}
    {Alex}{Tom}{J\"{u}rgen}
}
I have {\nToksFor{myPals}} pals, they are \useRanTok{1}, \useRanTok{2},
\useRanTok{3}, \useRanTok{4}, {\useRanTok{5}} and \useRanTok{6}. (Listed
in the order of best friend to least best friend.)

Test of the \verb!\bRTVToks!/\verb!\eRTVToks! pair of commands and of the
\texttt{rtVW} environment.

\bRTVToks{myThoughts}%
\begin{rtVW}
\begin{minipage}[t]{.67\linewidth}
Roses are red and violets are blue,
I've forgotten the rest, have you too?
\end{minipage}
\end{rtVW}
\begin{rtVW}
\begin{minipage}[t]{.67\linewidth}
I gave up saying bad things like
\verb!$#%%%^*%^&#$@#! when I was just a teenager
\end{minipage}
\end{rtVW}
\begin{rtVW}
\begin{minipage}[t]{.67\linewidth}
I am a good guy, pass it on! The code for this last sentence is,
\begin{verbatim}
%#$% I am a good guy, pass it on! ^&*&^*
\end{verbatim}
How did that other stuff get in there?
\end{minipage}
\end{rtVW}
\eRTVToks
\begin{enumerate}
    \item \useRanTok{1}
    \item \useRanTok{2}
    \item \useRanTok{3}
\end{enumerate}

Use \verb!\useRTName! command when another list separates the current
position from the list you want to use. Here we want to use the list named
\texttt{myPals}, but since that definition, a new list named
\texttt{myThoughts} was declared.

\useRTName{myPals}
List of pals: \useRanTok{1}, \useRanTok{2}, \useRanTok{3}, \useRanTok{4},
\useRanTok{4}, \useRanTok{5}.

For mixing lists. it might be easier to use the optional parameter:
{\useRanTok[myPals]{1}} and \useRanTok[myThoughts]{1}.

The \verb!\rtTokByNum! can retrieve an item from the list, in its declared
order; eg, from the \texttt{myPals} list, the first and last are {\rtTokByNum[myPals]{1}}
and \rtTokByNum[myPals]{\nToksFor{myPals}}.

\end{document}
