This Font Software is licensed under the SIL Open Font
License, Version 1.1.  This license is copied below, and is
also available with a FAQ at:
\url{http://scripts.sil.org/OFL}.

\def\ofl{\textsc{ofl}}

\subsection*{Preamble}
\label{sil:preamble}

The goals of the Open Font License (\ofl) are to stimulate
worldwide development of collaborative font projects, to
support the font creation efforts of academic and linguistic
communities, and to provide a free and open framework in
which fonts may be shared and improved in partnership with
others.

The \ofl\ allows the licensed fonts to be used, studied,
modified and redistributed freely as long as they are not
sold by themselves. The fonts, including any derivative
works, can be bundled, embedded, redistributed and/or sold
with any software provided that any reserved names are not
used by derivative works. The fonts and derivatives,
however, cannot be released under any other type of license.
The requirement for fonts to remain under this license does
not apply to any document created using the fonts or their
derivatives.

\subsection*{Definitions}
\label{sil:definitions}

\def\silterm#1{\noindent``#1''}

\begin{description}
\item[\silterm{Font Software}] refers to the set of files released
by the Copyright Holder(s) under this license and clearly
marked as such. This may include source files, build scripts
and documentation.

\item[\silterm{Reserved Font Name}] refers to any names specified
as such after the copyright statement(s).

\item[\silterm{Original Version}] refers to the collection of Font
Software components as distributed by the Copyright
Holder(s).

\item[\silterm{Modified Version}] refers to any derivative made by
adding to, deleting, or substituting --- in part or in whole
--- any of the components of the Original Version, by
changing formats or by porting the Font Software to a new
environment.

\item[\silterm{Author}] refers to any designer, engineer,
programmer, technical writer or other person who contributed
to the Font Software.
\end{description}

\subsection*{Permission \& Conditions}
\label{sil:permission}

Permission is hereby granted, free of charge, to any person
obtaining a copy of the Font Software, to use, study, copy,
merge, embed, modify, redistribute, and sell modified and
unmodified copies of the Font Software, subject to the
following conditions:

\begin{enumerate}
\item Neither the Font Software nor any of its individual
components, in Original or Modified Versions, may be sold by
itself.

\item Original or Modified Versions of the Font Software may
be bundled, redistributed and/or sold with any software,
provided that each copy contains the above copyright notice
and this license. These can be included either as
stand-alone text files, human-readable headers or in the
appropriate machine-readable metadata fields within text or
binary files as long as those fields can be easily viewed by
the user.

\item No Modified Version of the Font Software may use the
Reserved Font Name(s) unless explicit written permission is
granted by the corresponding Copyright Holder. This
restriction only applies to the primary font name as
presented to the users.

\item The name(s) of the Copyright Holder(s) or the
Author(s) of the Font Software shall not be used to promote,
endorse or advertise any Modified Version, except to
acknowledge the contribution(s) of the Copyright Holder(s)
and the Author(s) or with their explicit written permission.

\item The Font Software, modified or unmodified, in part or
in whole, must be distributed entirely under this license,
and must not be distributed under any other license. The
requirement for fonts to remain under this license does not
apply to any document created using the Font Software.
\end{enumerate}

\subsection*{Termination}
\label{sil:termination}

This license becomes null and void if any of the above
conditions are not met.

\subsection*{Disclaimer}
\label{sil:disclaimer}

\textsc{The Font Software is provided ``as is'', without
warranty of any kind, express or implied, including but not
limited to any warranties of merchantability, fitness for a
particular purpose and noninfringement of copyright, patent,
trademark, or other right. In no event shall the copyright
holder be liable for any claim, damages or other liability,
including any general, special, indirect, incidental, or
consequential damages, whether in an action of contract,
tort or otherwise, arising from, out of the use or inability
to use the Font Software or from other dealings in the Font
Software.}
