\documentclass{article}

\usepackage{starfont}
\usepackage{url}
\usepackage{hyperref}

\newcommand{\showboth}[1]{\starfontsans #1 & \starfontserif #1}

\begin{document}
\title{The \texttt{starfont} package for \LaTeX, version 1.2}
\author{Matthew Skala\thanks{\url{mailto:mskala@ansuz.sooke.bc.ca}}}
\date{29 September 2010}
\maketitle

\section{Introduction}

This package is designed to provide two fonts named StarFont Sans and
StarFont Serif in \LaTeX.  These fonts contain a wide variety of glyphs
suitable for typesetting astrological documents.  Astrological support in
previous \LaTeX\ packages is inadequate for a number of reasons that include
incomplete coverage of the popular glyph set, inconsistent style between
glyphs that need to work well when used together, simply ugly glyphs, wide
variation in widths of glyphs (which makes trouble in tables), no
availability in outline form to make well-behaved PDFs, and so on.  These
glyphs aren't perfect either, but they are certainly the best ones I've
found for typesetting astrological material in \LaTeX.

Given the typical use of \LaTeX\ in the scientific community and the
relationship between the scientific community and astrology, it amuses me to
have \LaTeX\ used for astrology.  It's also convenient because I write both
astrological and scientific documents and want to use the same tools for
both.

This is version 1.2 of the package, adding support for StarFont Serif and
changing the commentary on copying restrictions now that it's confirmed the
fonts are public domain.  Version 1.1 dated from 29~June 2006; version 1.0
was never widely distributed but first existed around 2003 or 2004.

The fonts were designed, and distributed in TrueType form, by
Anthony~I.P. Owen.  The packaging for TeX and LaTeX was done by Matthew
Skala.  Both these authors have released their contributions to the public
domain.  As such, the contents of this package are provided ``as is,'' with
no warranties.  See the file \texttt{COPYING} for more information.

\section{Using \texttt{starfont}}

Install the fonts.  Installing fonts is a black art; I have included, in the
\texttt{README} file, hints on where to put the files involved, but I do not
really understand how \LaTeX\ (and \TeX, and \texttt{dvips}, and so on)
finds font files and I cannot give meaningful instruction to others on it.

The fonts should end up installed with the names ``\texttt{fstr8x}'' for
StarFont Sans, and `\texttt{fsts8x}'' for StarFont Serif.  The
\texttt{starfont} package defines NFSS families named ``\texttt{stf}'' and
`\texttt{sts}''. The general approach used imitates the way the
\texttt{marvosym} package works, and this package is to some degree
compatible with \texttt{marvosym}.

StarFont Sans is the default and recommended choice; I believe it mixes
better with other fonts typically used in \LaTeX\ documents. However, the
StarFont Serif version is also available via the \texttt{starfontserif}
package option.  A corresponding \texttt{starfontsans} option will be
accepted for clarity but has no actual effect because the Sans version is
default anyway.  These options have verbose names to make it easier for them
to inherit from a higher-level package like my \texttt{horoscop} package. 
You can also switch between the two styles inside a document using the
\verb|\starfontsans| and \verb|\starfontserif| macros; each of these will
change the mode for any subsequent uses of the character macros below.  Both
are shown in the sample tables below.

\subsection{Luminaries and traditional planets}

Here are the basics:

\begin{tabular}{llll|llll}
\verb|\Sun| & \showboth{\Sun} & Sun &
   \verb|\Jupiter| & \showboth{\Jupiter} & Jupiter \\
\verb|\Moon| & \showboth{\Moon} & Moon &
   \verb|\Saturn| & \showboth{\Saturn} & Saturn \\
\verb|\Mercury| & \showboth{\Mercury} & Mercury &
   \verb|\Uranus| & \showboth{\Uranus} & Uranus \\
\verb|\Venus| & \showboth{\Venus} & Venus &
   \verb|\Neptune| & \showboth{\Neptune} & Neptune \\
\verb|\Terra| & \showboth{\Terra} & Earth (planet) &
   \verb|\Pluto| & \showboth{\Pluto} & Pluto \\
\verb|\Mars| & \showboth{\Mars} & Mars
\end{tabular}

The planet Earth is referred to as ``Terra'' in the command name to
prevent a collision with the glyph for the element Earth.  Since the
element is much more commonly referred to in astrology than the
planet, it takes precedence for the more straightforward name.  There
are also variants of some of these glyphs:

\begin{tabular}{llll|llll}
\verb|\varMoon| & \showboth{\varMoon} & Moon &
   \verb|\varUranus| & \showboth{\varUranus} & Uranus \\
\verb|\varTerra| & \showboth{\varTerra} & Earth (planet) &
   \verb|\varPluto| & \showboth{\varPluto} & Pluto
\end{tabular}

\subsection{Signs of the zodiac}

\begin{tabular}{llll|llll}
\verb|\Aries| & \showboth{\Aries} & Aries &
   \verb|\Libra| & \showboth{\Libra} & Libra \\
\verb|\Taurus| & \showboth{\Taurus} & Taurus &
   \verb|\Scorpio| & \showboth{\Scorpio} & Scorpio \\
\verb|\Gemini| & \showboth{\Gemini} & Gemini &
   \verb|\Sagittarius| & \showboth{\Sagittarius} & Sagittarius \\
\verb|\Cancer| & \showboth{\Cancer} & Cancer &
   \verb|\Capricorn| & \showboth{\Capricorn} & Capricorn \\
\verb|\Leo| & \showboth{\Leo} & Leo &
   \verb|\Aquarius| & \showboth{\Aquarius} & Aquarius \\
\verb|\Virgo| & \showboth{\Virgo} & Virgo &
   \verb|\Pisces| & \showboth{\Pisces} & Pisces
\end{tabular}

The command
\verb|\varCapricorn| produces ``\varCapricorn'' or
``\starfontserif\varCapricorn,'' a variant symbol for
Capricorn.  This package also
provides a \verb|\Zodiac| command similar to the one in \texttt{marvosym}:
\verb|\Zodiac{1}|\ldots\verb|\Zodiac{12}| produce
\starfontsans\Zodiac{1}\ldots\Zodiac{12} or
\starfontserif\Zodiac{1}\ldots\Zodiac{12}.

\subsection{Asteroids}

\begin{tabular}{llll|llll}
\verb|\Ceres| & \showboth{\Ceres} & Ceres &
   \verb|\Amor| & \showboth{\Amor} & Amor \\
\verb|\Pallas| & \showboth{\Pallas} & Pallas &
   \verb|\Eros| & \showboth{\Eros} & Eros \\
\verb|\Juno| & \showboth{\Juno} & Juno &
   \verb|\Hidalgo| & \showboth{\Hidalgo} & Hidalgo \\
\verb|\Vesta| & \showboth{\Vesta} & Vesta &
   \verb|\Hygiea| & \showboth{\Hygiea} & Hygiea \\
\verb|\Chiron| & \showboth{\Chiron} & Chiron &
   \verb|\Psyche| & \showboth{\Psyche} & Psyche \\
& & & & \verb|\Sappho| & \showboth{\Sappho} & Sappho
\end{tabular}

\subsection{Uranian hypothetical planets}

\begin{tabular}{llll|llll}
\verb|\Cupido| & \showboth{\Cupido} & Cupido &
   \verb|\Apollon| & \showboth{\Apollon} & Apollon \\
\verb|\Hades| & \showboth{\Hades} & Hades &
   \verb|\Admetos| & \showboth{\Admetos} & Admetos \\
\verb|\Zeus| & \showboth{\Zeus} & Zeus &
   \verb|\Vulkanus| & \showboth{\Vulkanus} & Vulkanus \\
\verb|\Kronos| & \showboth{\Kronos} & Kronos &
   \verb|\Poseidon| & \showboth{\Poseidon} & Poseidon
\end{tabular}

\subsection{Derived points}

\begin{tabular}{llll|llll}
\verb|\NorthNode| & \showboth{\NorthNode} & North~Node &
   \verb|\SouthNode| & \showboth{\SouthNode} & Sout~Node \\
\verb|\Lilith| & \showboth{\Lilith} & Lilith &
   \verb|\Fortune| & \showboth{\Fortune} & Fortune
\end{tabular}

\subsection{Aspects}

\begin{tabular}{llll|llll}
\verb|\Conjunction| & \showboth{\Conjunction} & Conjunction &
   \verb|\Quincunx| & \showboth{\Quincunx} & Quincunx \\
\verb|\Opposition| & \showboth{\Opposition} & Opposition &
   \verb|\Semisextile| & \showboth{\Semisextile} & Semisextile \\
\verb|\Trine| & \showboth{\Trine} & Trine &
   \verb|\Semisquare| & \showboth{\Semisquare} & Semisquare \\
\verb|\Square| & \showboth{\Square} & Square &
   \verb|\Sesquiquadrate| & \showboth{\Sesquiquadrate} & Sesquiquadrate \\
\verb|\Sextile| & \showboth{\Sextile} & Sextile
\end{tabular}

\subsection{Angles}

\begin{tabular}{llll|llll}
\verb|\ASC| & \showboth{\ASC} & Ascendant &
   \verb|\DSC| & \showboth{\DSC} & Descendant \\
\verb|\MC| & \showboth{\MC} & Midheaven &
   \verb|\IC| & \showboth{\IC} & Nadir \\
 & & & (Medium C\oe li) & & & & (Imum C\oe li) \\
\verb|\Vertex| & \showboth{\Vertex} & Vertex &
   \verb|\EastPoint| & \showboth{\EastPoint} & East Point
\end{tabular}

\subsection{Directions}

\begin{tabular}{llll|llll}
\verb|\Retrograde| & \showboth{\Retrograde} & Retrograde &
   \verb|\Station| & \showboth{\Station} & Station \\
\verb|\Direct| & \showboth{\Direct} & Direct
\end{tabular}

\subsection{Elements}

\begin{tabular}{llll|llll}
\verb|\Fire| & \showboth{\Fire} & Fire &
   \verb|\Earth| & \showboth{\Earth} & Earth (element) \\
\verb|\Air| & \showboth{\Air} & Air &
   \verb|\Water| & \showboth{\Water} & Water
\end{tabular}

\subsection{Other symbols}

\begin{tabular}{llll|llll}
\verb|\Natal| & \showboth{\Natal} & Natal &
   \verb|\Radix| & \showboth{\Radix} & Radix \\
\verb|\Pentagram| & \showboth{\Pentagram}
\end{tabular}

There are many other symbols in the font which I don't recognize,
including a bunch that are probably from Cosmobiology.  It should be
possible to add them easily enough by following the examples in
\texttt{starfont.sty}. There is an encoding table (made by
\texttt{testfont.tex}) provided in the file \texttt{table.pdf}.  I
would appreciate hearing from readers who know of good names for the
unknown symbols.

\end{document}
