% This file is public domain.
% Originally written 1998, Ivan A Derzhanski.
% Notice added by Clea F. Rees 2009/01/06.
\documentclass{article}
\usepackage{a4wide}
\pagestyle{empty}

\font\cmdh=cmdunh10   \font\eiaddh=eiaddunh10
\font\cmfib=cmfib8    \font\eiadfib=eiadfib8
\font\cmvtt=cmvtt10   \font\eiadvtt=eiadvtt10

\begin{document}
\newcommand\english{\fontfamily{cmr}\selectfont}
\newcommand\irish{\fontfamily{eiad}\selectfont}
\irish Amhra/n macaro/nach a chuir {\english Marion Gunn}
chuig \textsc{\english gaelic-l}.
Le/iri/onn an te/acs seo aicme na gclo/nna Gaelach rialta
chomh maith le foirinn chlo/scri/ofa an leithid luainigh,
\english Dunhill {\irish7`} Fibonacci.
(This text demonstrates the Regular Irish family of typefaces
as well as the variable-width typewriter fount, Dunhill and Fibonacci.)

\begin{verse}

{\fontseries{b}\english One day for recreation \quad
\irish Is gan e/inne beo im chuideachta, \\
\english I spied a charming fair maid, \quad
\irish Ina haonar is i/ i siopa istigh.} \\
{\bfseries\english She was singing like an angel, \quad
\irish Is me/ ag e/isteacht lena binne-ghuth; \\
\english I whispered soft and easy--- \quad
\irish Is e/ du/irt si/: Stad ded radaireacht!}

\qquad \irish Curfa//\english Chorus:
         
\irish \qquad\qquad Anonn is anall, a Mha/iri/n, \quad
       \quad Do mha/lai/ is do bheilteanna; \\
       \qquad\qquad Is a bhean na stocai/ mba/na, \quad
       \quad Ba bhrea/ liom bheith ag iomaidh leat.

\slshape

\english Her amber locks most neatly \quad
\irish Go dre/imreach si/os ag titim le/i, \\
\english Adown her back and waist, \quad
\irish Is gur phreab mo chroi/ le taitneamh di. \\
{\bfseries\english I asked her was she the fair one, \quad
\irish An bhandia u/d bhi/ ag Iu/patar \\
\english Or the brightsome vestal deity \quad
\irish A chaith tre/imhse seal in Ifreann.}

\qquad Curfa//\english Chorus.
         
\itshape

\english She answered me most daintily: \quad
\irish Ni/ he/inne da/r thugais me/. \\
\english I fear you are a \irish re/ice, \quad
Is na/ taobhaigh a thuilleadh me/! \\
{\bfseries\english Indeed I am no \irish re/ice, \quad
Na/ straeire a bhre/agfadh bruinneall seal. \\
\english I'm a pupil of Jack Lahey's; \quad
\irish Is is i/ an a/it a gco/nai/m Mucros.}

\qquad Curfa//\english Chorus.
         
\upshape

{\scshape\english I asked her who her father was, \quad
\irish Is du/irt si/ liom an ministir. \\
\english I knew I stood in danger, \quad
\irish Is gur bhaolach da/ bhfeicfi/ sinn!} \\
\let\english\cmvtt   \let\irish\eiadvtt
\english If I had you in a neat grove, \quad
\irish Idir Claonach is Mucros, \\
\english Your sparkling eyes do tease me, \quad
\irish Tri/ la/r mo chroi/ ta/ taitneamh duit.

\qquad Curfa//\english Chorus.
         
\let\english\cmdh   \let\irish\eiaddh
\english Her syllables most charming, \quad
\irish Is gur bhrea/ liom bheith ina cuideachta. \\
Is gur bhinne liom na/ an chla/irseach \quad
Gach ardphort da/ seinneadh si/. \\
\let\english\cmfib   \let\irish\eiadfib
\english You'll get my stock and farm, \quad
\irish Ma/ the/ann tu/ liom go Mucros. \\
\english And then she sang most charming: \quad
\irish A ghra/ geal, \english I'm fond of you!

\qquad \irish Curfa//\english Chorus.
\end{verse}
\end{document}

