%%  (c) copyright  2006, 2007
%% Antonis Tsolomitis
%% Department of Mathematics, University of the Aegean
%%
%%  This document can be redistributed and/or modified under the terms
%%  of the LaTeX Project Public License Distributed from CTAN
%%  archives in directory macros/latex/base/lppl.txt; either
%%  version 1 of the License, or any later version.

\documentclass{article}
\usepackage[polutonikogreek,english]{babel}
\usepackage[iso-8859-7]{inputenc}
\usepackage{cmbright}
%\usepackage[default]{gfsneohellenic}
%\usepackage{gfsneohellenic}
\usepackage{gfscomplutum}
\usepackage{latexsym,amsfonts}
%\renewcommand{\ttdefault}{hlst}

%%%%% Theorems and friends
\newtheorem{theorem}{Θεώρημα}[section]         
\newtheorem{lemma}[theorem]{Λήμμα}
\newtheorem{proposition}[theorem]{Πρόταση}
\newtheorem{corollary}[theorem]{Πόρισμα}
\newtheorem{definition}[theorem]{Ορισμός}
\newtheorem{remark}[theorem]{Παρατήρηση}
\newtheorem{axiom}[theorem]{Αξίωμα}
\newtheorem{exercise}[theorem]{¶σκηση}


%%%%% Environment ``proof''
\newenvironment{proof}[1]{{\textit{Απόδειξη:}}}{\ \hfill$\Box$}
\newenvironment{hint}[1]{{\textit{Υπόδειξη:}}}{\ \hfill$\Box$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%






\title{The \textsc{gfs complutum} font}
\author{Antonis Tsolomitis\\
Laboratory of Digital Typography\\ and Mathematical Software\\
Department of Mathematics\\
University of the Aegean\\
\textlatin{\texttt{http://myria.math.aegean.gr/labs/dt}}}
\date {\textsc{12} May \textsc{2006}}


\begin{document}
\maketitle

\section{Introduction}



The font---and hence the package---supports only the LGR encoding and it 
is to be used with the greek option of the Babel package.

The fonts are loaded with

\verb|\usepackage{gfscomplutum}| or \verb|\usepackage[default]{gfscomplutum}|.


The latter one sets \textsc{gfs} Complutum as the default document font. The former
defines the environment \verb|complutum| and the command
\verb|\textcomplutum|. For example, while in Greek language

\verb|\textcomplutum{dokim'h}| 

\noindent produces 

\textgreek{\textcomplutum{dokim'h}}.

x-height is adjusted so that it matches with the x-height of the
\verb|gfsneohelenic| package. 




\section{Installation}

Copy the contents of the subdirectory afm in
texmf/fonts/afm/GFS/Complutum

\medskip

\noindent Copy the contents of the subdirectory doc in
texmf/doc/latex/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory enc in
texmf/fonts/enc/dvips/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory map in
texmf/fonts/map/dvips/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory tex in
texmf/tex/latex/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory tfm in
texmf/fonts/tfm/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory type1 in
texmf/fonts/type1/GFS/Complutum/

\medskip

\noindent Copy the contents of the subdirectory vf in
texmf/fonts/vf/GFS/Complutum/

\medskip

\noindent In your installation's updmap.cfg file add the line

\medskip

\noindent Map gfscomplutum.map

\medskip

Refresh your filename database and the map file database (for example, on Unix systems
run mktexlsr and then run the updmap-sys (or updmap on older systems) script as root).

You are now ready to use the fonts.

\section{Usage}

As said in the introduction the package covers only
greek. Greek covers polytonic too through babel (read the
documentation
of the babel package and its greek option). 

For example, the preample

\begin{verbatim}
\documentclass{article}
\usepackage[english,greek]{babel}
\usepackage[iso-8859-7]{inputenc}
\usepackage{cmbright}
\usepackage[default]{gfscomplutum}
\end{verbatim}

will be the correct setup for articles in Greek using Complutum for the
main font.

\bigskip

\subsection{Transformations by \texttt{dvips}}

Other than the shape provided by the font itself, this package
provides a slanted shape as well
using the standard mechanism provided by dvips. 


\subsection{Alternative characters}
There is an alternative design for tau. This is used by default as the
second of two consecutive taus, as for example in
\textgreek{\textcomplutum{Αττική}} (Attika).
This form of tau can be accessed directly with \verb|\taualt|.




\section{Problems}


 The
accents of the capital letters should hang in the left margin when such a letter starts a
line.  \TeX\ and \LaTeX\ do not provide the tools for such a
feature. However, this seems to be possible with 
\textlatin{pdf\TeX}
As this is work in progress, please be patient\ldots




\section{Samples}


\Large
\greektext\selectlanguage{polutonikogreek}
\begin{complutum}
\begin{center}
<'Hrhc
\end{center}
\begin{flushright}
θυμίαμα αρώματα
\end{flushright}
\begin{center}
%\begin{minipage}{10cm}
\begin{verse}
Κυανέοις κόλποισιν >ενημένη, >αερόμορφε,\\
<'Ηρα παμβασίλεια, Δι`ος σύλλεκτρε μάκαιρα,\\
ψυχοτρόφους a>'urac θνητο~ις παρέχουσα προσηνε~ις,\\
>'ομβρων μ`εν μήτηρ, >ανέμων τροφέ, παντογένεθλε:\\
χωρ`ις γ`αρ σέθεν ο>υδ`εν <'ολως ζω~ης φύσιν >'εγνω:\\
κοινωνε~ις γ`αρ <'απασι κεκραμένη >ηέρι σεμν~ωι:\\
πάντων γ`αρ κρατέεις μούνη πάντεσσί τ'' >ανάσσεις\\
>ηερίοις <ροίζοισι τινασσομένη κατ`α χε~υμα.\\
>αλλά, μάκαιρα θεά, πολυώνυμε, παμβασίλεια,\\
>'ελθοις ε>υμενέουσα καλ~ωι γήθοντι προσώπωι.\\
\end{verse}
%\end{minipage}
\end{center}
\end{complutum}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
